A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-17 papers/static-checker/ecoop2024/ecoop2024.tex
update some specifications

Our starting point is the design of a domain-specific language for
specifying smart contract properties.
\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification (excerpt)},label={lst:auction-fragment}]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                  Pair (highest_bidder: address)
                       (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open = true
       && Amount > pre_Balance 
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, pre_Balance, c) ;
\end{lstlisting}
As an example, Listing~\ref{lst:auction-fragment} contains an excerpt
of the specification of an auction contract. The \texttt{storage}
phrase describes the internal state (storage) of the contract
consisting of three values named \lstinline|auction_open|,
\lstinline|highest_bidder|, and \lstinline|contract_owner|.
The contract has several entrypoints that may be invoked individually.

@@

\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax (option 1)},label={lst:multiple-entrypoint-specification-1},numbers=left][h]
mcontract Name = spec
   entrypoint %a 
      code := {c1}
      input := t1
      output := t1'
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      ...
   <entrypoint relations>
\end{lstlisting}

\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax (option 2)},label={lst:multiple-entrypoint-specification-2},numbers=left][h]
mcontract Name = spec
   storage := t1
   entrypoint %a 
      code := {c1}
      parameter := t2
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      ...
   <entrypoint relations>
\end{lstlisting}

@@

  (%create -> %bid) with (auction_open = true)  
       and (Amount > pre_Balance) 
       and get_contract (unit, highest_bidder) = 
           Some (c: contract unit)
  | (%create  -> %close) with (auction_open = true) 
       and (Sender = contract_owner)
       and get_contract (unit, contract_owner) = 
          Some (c: contract unit)
  | (%bid -> %bid) with (Amount > Balance - Amount) 
       and get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
  | (%bid -> %close) with (Sender = contract_owner) 
       and get_contract (unit, contract_owner) = 
            Some (c: contract unit)
  | not (%close -> %bid)
  | not (%close -> %close)
\end{lstlisting}

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}\footnote{There may be some duplicates of the same failwith condition due to the branching of the symbolic interpreter. The results here are shortened. We are working on filtering them in the tool.}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) &&
   get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   sbvar1 < value : 'NotEnoughBalance'
4. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (Sender, sbvar2) = Some sbvar3 &&
   sbvar3 < value: 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
 storage :=  Pair (ledger: big_map address 
  (pair nat (map address nat))) 
  (Pair (admin: address) 
          (Pair (paused: bool) (total_supply: nat)))
 entrypoint %transfer
   parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
   pre-condition := not (paused  = true)
     && get_big_map (from, ledger) = 
       Some (v: pair nat (map address nat))
     && (v = Pair (n: nat) (m: map address nat))
     && get_map (Sender, m) = Some (k: nat) 
     && (n >= value) && (k >= value)
   post-condition := get_big_map (from, post_ledger) = 
     Some (post_v:  pair nat (map address nat))
     && post_v = Pair (post_n: nat) (post_m: map address nat) 
     && get_map (Sender, post_m) = Some (post_k: nat) 
     && (post_n = n - value) && (post_k = k - value)
\end{lstlisting}

@@

&& (Now - t >= 0) && (Now - t >= maxDelay): 17

>>>>

Our starting point is the design of a domain-specific language for
specifying smart contract properties.
\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification (excerpt)},label={lst:auction-fragment}]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                  Pair (highest_bidder: address)
                       (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open = true
       && Amount > pre_Balance 
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open = true
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, pre_Balance, c) ;
\end{lstlisting}
As an example, Listing~\ref{lst:auction-fragment} contains an excerpt
of the specification of an auction contract. The \texttt{storage}
phrase describes the internal state (storage) of the contract
consisting of three values named \lstinline|auction_open|,
\lstinline|highest_bidder|, and \lstinline|contract_owner|.
The contract has several entrypoints that may be invoked individually.

@@

\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax (option 1)},label={lst:multiple-entrypoint-specification-1},numbers=left][h]
mcontract name = spec
   entrypoint %a 
      code := {c1}
      input := t1
      output := t1'
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      ...
   <entrypoint relations>
\end{lstlisting}

\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax (option 2)},label={lst:multiple-entrypoint-specification-2},numbers=left][h]
mcontract name = spec
   storage := t1
   entrypoint %a 
      code := {c1}
      parameter := t2
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      ...
   <entrypoint relations>
\end{lstlisting}

@@

  (%create -> %bid) with (auction_open = true)  
       and (Amount > pre_Balance) 
       and get_contract (unit, highest_bidder) = 
           Some (c: contract unit)
  | (%create  -> %close) with (auction_open = true) 
       and (Sender = contract_owner)
       and get_contract (unit, contract_owner) = 
          Some (c: contract unit)
  | (%bid -> %bid) with (Amount > pre_Balance) 
       and get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
  | (%bid -> %close) with (Sender = contract_owner) 
       and get_contract (unit, contract_owner) = 
            Some (c: contract unit)
  | not (%close -> %bid)
  | not (%close -> %close)
\end{lstlisting}

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}\footnote{There may be some duplicates of the same failwith condition due to the branching of the symbolic interpreter. The results here are shortened. We are working on filtering them in the tool.}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) &&
   get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   sbvar1 < value : 'NotEnoughBalance'
4. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (Sender, sbvar2) = Some sbvar3 &&
   sbvar3 < value: 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
 storage :=  Pair (ledger: big_map address 
  (pair nat (map address nat))) 
  (Pair (admin: address) 
          (Pair (paused: bool) (total_supply: nat)))
 entrypoint %transfer
   parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
   pre-condition := not (paused  = true)
     && get_big_map (from, ledger) = 
       Some (v: pair nat (map address nat))
     && v = Pair (n: nat) (m: map address nat)
     && get_map (Sender, m) = Some (k: nat) 
     && (n >= value) && (k >= value)
   post-condition := get_big_map (from, post_ledger) = 
     Some (post_v:  pair nat (map address nat))
     && post_v = Pair (post_n: nat) (post_m: map address nat) 
     && get_map (Sender, post_m) = Some (post_k: nat) 
     && (post_n = n - value) && (post_k = k - value)
\end{lstlisting}

@@

&& (Now - t >= 0) && not (Now - t < maxDelay): 17
